---
title: "Patterns across players in champion and role selection"
output: html_document

---

```{r setup, include=FALSE}
library(plyr) #tidyverse
library(dplyr)#tidyverse
library(tidyverse)#tidyverse
library(ggplot2) #plotting
library(ggdark) #plotting
library(reshape2) #for melting into matrix

```

## The data

First I read in match-level data which I had saved as separate 5 separate .csvs. 

I curated this data by requesting data from Riot's public API (https://developer.riotgames.com/) using R (see file in repo fetch_matches_riot_api.R). The logic was as follows:

1) Request summoner IDs of 1800 summoners (Samples 300 in each rank from Iron to Diamond, 75 per division)

2) Using summoner IDs, I then requested puuids (another form of account ID which is needed to pull match data )

3) Use puuids to request a list of the last 30 ranked matches for each summoner as match IDs

4) Using match IDs I accessed the match-level data of for all these matches

Together this gives me information about ~1800 summoners most recent 30 ranked games which could let track player patterns such as the following:

Some examples of questions you could answer with this dataset
What roles does a player usually play?
If a players plays one lane, what other lanes do they tend to play?
How many champions do summoners tend to play (in their last 30 games)? Are most players one-trick ponies or do they have a wide champion pool? Does this differ by lane?
How many roles do summoners tend to play?
After a win, do summoners play the same champion again?

Ideally it would be great to track a player's match data across an entire season. Unfortunately this would take a long time due to limited number of requests I am allowed to make on the API. Still looking at past 30 ranked matches, games players are likely taking serious, can still give some important insights of player patterns and what roles they tend to play! There is also a neat Kaggle dataset (which inspired this project) that tracked Gold-level players across an entire ranked season, which now a bit outdated, can give some really rich insights into how player behavior changes across a season.

```{r wrangling, echo = FALSE}

#Read in all .csvs containing match-level (I have it separated in 6 batches)
setwd("~/Documents/GitHub/lol/lol_analysis/match_data")
match_files <- list.files(recursive=TRUE, pattern = "*match")
matches  <-do.call(rbind.fill,lapply(match_files,read.csv))   %>%
  filter(teamPosition != '') #Remove matches when we don't have position information (empty string)

#How many summoners are we dealing with (there were some summoners who we had some trouble retrieving puuids for)
#length(unique(matches$puuid))

#Lets see how much data we got for each summoner
sum_level <- filter(matches, teamPosition != '') %>%
  group_by(puuid) %>%
  summarize(count = length(puuid))

#How many summoners do we have more than 30 recent matches from
print('number of players we have more than 30 recent ranked games')
sum_level %>%
  filter(count > 29) %>%
  nrow()

#Lets see how much data we got for each summoner
sum_level <- matches %>%
  group_by(puuid) %>%
  summarize(count = length(puuid))


#These are summoners that don't repeat across a lot of matches
sum_no_rep <- sum_level$puuid[which(sum_level$count < 10)]

#We want to get the 30 most recent games 
sum_30 <- sum_level %>%
  filter(count >29)

#List of summoners with more than 30 games

#Go through each summoner and label their games
match_sort <- filter(matches, puuid %in% sum_30$puuid) %>%
  arrange(puuid, game_start) #Arrange by summoner and time (unix time)
  
  
#Make new column (match_order) which gives a number to each match where 1 = earliest game and higher numbers = later games
match_sort$match_order <- NA
match_counter = 1
for (i in 1:nrow(match_sort)) {

  
  #if new summoner reset counter
  if ((i != 1) && (match_sort$puuid[i] != match_sort$puuid[i-1])) {
    match_counter <- 1
    
  }
  
  match_sort$match_order[i] <- match_counter
  
  match_counter <- match_counter + 1
  
  
  
} 

#Now to choose first 30 games for each summoner can exclude any match_order values above 30
match_sort <- filter(match_sort, match_order < 31)


```




## Unique champion analysis
For each summoner I took their 30 most recent ranked games (solo queue) and looked at how many different (unique) champions they played. There is quite a bit of variability in how many champions people play. Most summoners play somewhere between 5-10 champions (at least in their last 30 games)

```{r unique_champions, echo = FALSE, fig.height = 6, fig.width = 12, message = FALSE, warning = FALSE}
#How about number of unique champions being played per person?
unique_champions<- filter(match_sort) %>%
  group_by(puuid) %>%
  summarize(unique_champs_played = length(unique(championName)),
            total_games = n())

#Mean number of champs played
grp.mean = mean(unique_champions$unique_champs_played)
grp.median = median(unique_champions$unique_champs_played)

ggplot(unique_champions, aes(x=unique_champs_played))+  
  geom_histogram(aes(y = (..count..)/sum(..count..)), binwidth = 1, color = 'darkblue', fill = 'lightblue')+ 
  dark_theme_gray(base_size = 20, base_family = 'Helvetica') +
  #stat_bin(aes(y=(..count..)/sum(..count..)*100, 
     #        $ label=paste0(round((..count..)/sum(..count..)*100,1),"%")),color = 'darkblue', fill = $'lightblue', bins = 30) +
  geom_vline(data=unique_champions, aes(xintercept=grp.mean),
             color="lightblue",
             linetype="dashed") +
  geom_vline(data=unique_champions, aes(xintercept=30),color="#d5f5b8") +
  geom_vline(data=unique_champions, aes(xintercept=0),color="#d5f5b8") +
  geom_text(aes(x=11.5, label="Average", y=0.1), colour="lightblue", angle=0, size = 6)+ 
  geom_text(aes(x=29, label="Different every game", y=0.072), colour="#d5f5b8", angle=90, size = 6) + 
  geom_text(aes(x=-1, label="Same every game", y=0.07), colour="#d5f5b8", angle=90, size = 6) +
  #coord_cartesian(ylim = c(0,0.12)) +
  ggtitle('How many champions do summoners play?') +
  theme(plot.background = element_rect(fill = "grey10"),
                panel.background = element_blank(),
                panel.grid.major = element_line(color = "grey30", size = 0.1),
                panel.grid.minor = element_line(color = "grey30", size = 0.1),
                legend.background = element_blank(),
                axis.ticks = element_blank(),
        legend.position = 'none')  +
  scale_x_continuous(breaks = seq(0, 30, by = 5))+
  labs(y = 'Proportion of players', x = "Unique champions played in last 30 ranked games")

#Narrow champ pool (Exploit, Mastery) ---> Wider champ pool (Explore, Flexibility)








```





## Unique role analysis
I also did the exact same thing as above except at the level of roles (i.e. top, mid, adc etc.). Most summoners seem to play three unique roles. Though 10% of players only played a single role (i.e. strong main). Again this is only in their past 30 ranked games.

```{r unique_roles, echo = FALSE, fig.height = 6, fig.width = 12}

#How many unique roles do summoners player ?
#Sometimes we don't get good role data (teamPosition is null) so we want to exclude these
unique_roles<- filter(match_sort, teamPosition != '') %>% 
  group_by(puuid) %>%
  summarize(unique_roles_played = length(unique(teamPosition)),
            total_games = n())  #Lets only look at players we have a lot of data on

#Mean number of roles played
grp.mean = mean(unique_roles$unique_roles_played)

ggplot(unique_roles, aes(x=unique_roles_played))+  
  geom_histogram(aes(y = (..count..)/sum(..count..)), binwidth = 1, color = 'darkblue', fill = 'lightblue')+ 
  dark_theme_gray(base_size = 20,base_family = 'Helvetica') +
  #stat_bin(aes(y=(..count..)/sum(..count..), 
              # label=paste0(round((..count..)/sum(..count..)*100,1),"%")),color = 'darkblue', fill = 'lightblue') +
  geom_vline(data=unique_roles, aes(xintercept=grp.mean, color="pink"),
             linetype="dashed") +
  #geom_vline(data=unique_roles, aes(xintercept=30, color="green")) +
  #geom_vline(data=unique_roles, aes(xintercept=0, color="green")) +
  #geom_text(aes(x=1, label="10% of players played the same lane for the past 30 games", y=0.27), colour="lightblue", angle=0, size = 4)+
  #geom_text(aes(x=29, label="Different every game", y=0.072), colour="pink", angle=90, size = 6) + 
  #geom_text(aes(x=-1, label="Same every game", y=0.07), colour="pink", angle=90, size = 6) +
  coord_cartesian(ylim = c(0,0.3))  +
  scale_x_continuous(breaks = seq(0, 5, by = 1)) +
  ggtitle('How many different roles do summoners play?') +
  theme(plot.background = element_rect(fill = "grey10"),
                panel.background = element_blank(),
                panel.grid.major = element_line(color = "grey30", size = 0.1),
                panel.grid.minor = element_line(color = "grey30", size = 0.1),
                legend.background = element_blank(),
                axis.ticks = element_blank(),
        legend.position = 'none') +
  labs(y = 'Percentage of players', x = "Unique roles played in last 30 ranked games")


```



## Are there patterns in what roles people tend to play?

```{r role_patterns2, echo = FALSE, fig.height = 4, fig.width = 6}


#First we want to identify what roles people main by looking at the percentage of games played in a certain position
role_mains <- match_sort %>% #Only look at summoners with more than 30 games worth of data
  group_by(puuid) %>%
  summarize(top_games = length(puuid[which(teamPosition == 'TOP')]),
         jung_games =length(puuid[which(teamPosition == 'JUNGLE')]),
         mid_games = length(puuid[which(teamPosition == 'MIDDLE')]),
         adc_games = length(puuid[which(teamPosition == 'BOTTOM')]),
         sup_games = length(puuid[which(teamPosition == 'UTILITY')]))%>%
  mutate(total_games = mid_games+top_games+jung_games+adc_games+sup_games, #Calculate percentage of games in each role
         mid_percent = mid_games/total_games,
         top_percent = top_games/total_games,
         jung_percent = jung_games/total_games,
         adc_percent = adc_games/total_games,
         sup_percent = sup_games/total_games)


#head(role_mains)


#Test correlation matrix old fashion way
r <- select(role_mains, top_games, mid_games, jung_games, sup_games, adc_games) %>%
  rename(Top = top_games, Jungle = jung_games, Mid = mid_games, Adc = adc_games, Support = sup_games )

#Mke correlatoin matrix
cor <- cor(r, method = "pearson", use = "complete.obs")
melted_cormat <- melt(cor)
clust <- hclust(dist(t(cor)))


#Clustered!!
ggplot(data = melted_cormat, aes(x=as.character(Var1),  y= as.character(Var2), fill=value)) + 
  geom_tile()+
  scale_fill_gradient2(low = "white", high = "darkred", mid = "darkred", 
                       midpoint = -0.15, limit = c(-0.35,-0.2), space = "Lab", 
                       name="")  + 
  dark_theme_gray(base_size = 18,base_family = 'Helvetica') + 
  scale_y_discrete(limits = colnames(cor)[clust$order])+
  scale_x_discrete(limits = colnames(cor)[clust$order]) +
  labs(y = '', x = '') 
  #ggtitle('What mains are tightly correlated?')



```



## If you main this role... what other roles do you play?)

```{r role_patterns, echo = FALSE, fig.height = 3, fig.width = 6}


role_name <- 'Jungle'
role_label <- 'jung_percent'
calc_role_mains <- function(role_name, role_label) {
  
  #Identify mains (main = the role with the most games)
  role_mains2 <- select(role_mains,top_percent, jung_percent, mid_percent, adc_percent, sup_percent) %>%
  mutate(main = names(.)[max.col(.)])
  role_mains2$puuid <- role_mains$puuid
  this_role <- filter(role_mains2, main ==  role_label)

this_role_long <- select(this_role, puuid, top_percent,mid_percent,jung_percent,adc_percent,sup_percent) %>%
  gather(role, percent, top_percent:sup_percent) 

this_role_long$role <- recode_factor(this_role_long$role, top_percent = "Top", 
                           jung_percent = "Jungle",
                           mid_percent = 'Mid',
                           adc_percent = 'ADC',
                           sup_percent = 'Support')

#Summarizze percent of each role
role_sum <- this_role_long %>%
  group_by(role) %>%
  summarise(mean_percent_games = mean(percent)) %>%
  filter(role != role_name) %>%
  arrange(desc(mean_percent_games)) #Reorder based on most to least played

#Reorder factors for plotting
role_sum$role <- factor(role_sum$role, levels=role_sum$role)

#But if its ranked its as expected because you need all roles?
role_p <- ggplot(role_sum, aes(x=role, y=mean_percent_games*100, fill=role)) +
  geom_bar(stat="identity")  + 
  dark_theme_gray(base_size = 18,base_family = 'Helvetica') +
  scale_fill_brewer(palette="Blues") +
  coord_cartesian(ylim = c(0,10)) +
  theme(legend.position = 'none')+
  labs(y = 'Percent of games', x = "Other roles") +
  ggtitle(paste('If you main ', role_name, sep = ''))

role_p



}

calc_role_mains('Top', 'top_percent')
calc_role_mains('Jungle', 'jung_percent')
calc_role_mains('Mid', 'mid_percent')
calc_role_mains('ADC', 'adc_percent')
calc_role_mains('Support', 'sup_percent')


```




